site_map

### 2019-09-23 SST_patterns.Rmd
progress

As a common date we will use a start date from 1992-01-31 and end date 17-01-31

>>>>

As a common date we will use a start date from 1992-01-31 and end date 16-12-31

```{r}
# Generalosed linear models
#Using the generalised linear hypotheses - Look at the difference in metrics (duration and intensity) of in situ upwelling and remotely sensed SST data upwelling

lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```

# From the years 1992 - 2017

```{r}
# Which pixels showed the highest upwelling counts within the different datasets
# Which site showed the most intense upwelling within the different datasets
# Where is the largest upwelling event occuring?

metric_func <- function(df){
 metrics <-df %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  #mutate(year = year(date_start)) %>% 
  group_by(distance, site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
   select(site, distance, mean_intensity, sum_events)
}

OISST_metrics <- metric_func(OISST_upwell_base)
CMC_metrics <- metric_func(CMC_upwell_base)

# No of u pwelling events found at the different sites
SACTN_metrics <- SACTN_upwell_base %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  group_by(site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
  select(site, mean_intensity, sum_events)

# Create a table showing these metrics

Create visualisation which compares the presence and absence of events
If a event is found in a particular site at a particulate dat, is the same event present in the other datasets? Show how the events detected varies between site and dataset.

```{r}
# To determine wether or not the same upwlling patterns are present at different distances from the coastline

- Pvalues
- plotting
- ERA5
- Plot wind data

### 2019-09-23 Write_up/Paper1_writeup.Rmd
progress

Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of the motion caused by alongshore equatorial winds. These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references). Occurring globally throughout the world's oceans, four major eastern boundary upwelling systems (EBUS) can be recognised; each significantly impacting associated coastal systems. These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly]. The inshore waters off the west coast of South Africa are dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews] (RWS: This could be made more clear. What is the main point you want to communicate in this sentence?). Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world. As such, there has been a constant interest in upwelling systems. Major upwelling programs (RWS: This is too vague. What are these programs?) have already been carried out in Peru [@walsh] and North-West Africa [@barber].

@@

Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this). These temperatures have increased as a result of anthropogenically driven climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed). This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. Climate change impacting oceanic processes has major ecological consequences [@walther]. This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance. Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters [@hoegh]. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper. Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions. (RWS: There appear to be two or three separate main points being made in this paragraph. I would recomend you split it up into a couple of paragraphs and think about what the main message of each should be.)

@@

>>>>

Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of the motion caused by alongshore equatorial winds. These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references). Occurring globally throughout the world's oceans, four major eastern boundary upwelling systems (EBUS) can be recognised; each significantly impacting associated coastal systems. These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly]. There has been significant interest in major upwelling systems since the later 1960s, with major interest occuring off Peru [@walsh] and North-West Africa [@barber]. Research done by Copenhagen 1953, reported a number of sites along the west coast of southern Africa where upwelling was particularly intense. The two sites of intense upwelling activity were found off the Cape Peninsula and off Cape columbine [@copenhagen; @andrew]

@@

Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this). These temperatures have increased as a result of anthropogenically driven climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed). This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. Climate change impacting oceanic processes has major ecological consequences [@walther]. This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance. Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters [@hoegh]. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper. Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions. (RWS: There appear to be two or three separate main points being made in this paragraph. I would recomend you use what you have here and try and work these main points into other paragraphs)

@@

(RWS: This is a very thorough introduction section. Impotantly see that they all flow from one paragraph to the next. If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)

### 2019-09-23 site_map.Rmd
progress

```{r}
load("Data/south_africa_coast.RData")
load("Data/africa_coast.RData")
load("Data/sa_provinces_new.RData")
load("Data/site_list_sub.Rdata")
load("Data/site_pixels.RData")
load("MUR.RData")
names(south_africa_coast)[1] <- "lon"

@@

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  xlab("") + ylab("") +
  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
  #            size = 4, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size =4, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 4, angle = 302, colour = "black") +
    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
    #          size = 5, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5),
        axis.text = element_text(size = 10)) +
    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.direction = "horizontal",
        legend.justification = c(1,0),
        legend.position = c(0.65, 0.80),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

@@

pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
print(site_map, vp = vp2)
print(africa_map, vp = vp1)
dev.off()

>>>>

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline. The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=6, fig.width=10, fig.pos="H"}

load("Data/south_africa_coast.RData")
load("Data/africa_coast.RData")
load("Data/sa_provinces_new.RData")
load("Data/site_list_sub.Rdata")
load("Data/site_pixels.RData")
load("MUR.RData")
names(south_africa_coast)[1] <- "lon"

@@

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + bw_update +
  geom_raster(data = MUR, aes(x = lon, y = lat, fill = temp)) +
   geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group),
                fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "white", shape = 0, alpha = 0.8, size = 2.1) +
  coord_equal(xlim = c(15, 25), ylim = c(-36,-26), expand = 0) +
  # geom_rect(data = bbox, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #           alpha = 0.4, colour = "red", size = 0.1, linetype = 1) +
  # scale_x_continuous(limits = sa_lons, expand = c(0, 0), breaks = seq(15, 35, 5)) +
  # scale_y_continuous(limits = sa_lats, expand = c(0, 0), breaks = seq(-35, -30, 5)) +
  scale_fill_viridis(
    option = "magma",
    name = "Temperature (°C)",
    alpha = 0.8) +
  xlab("") + ylab("") +
  # annotate("text", label = "INDIAN\nOCEAN", x = 32.00, y = -35.0,
  #            size = 4, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 17.00, y = -35.0,
             size =4, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 16.0, y = -31.7,
             size = 4, angle = 302, colour = "black") +
    # annotate("text", label = "Agulhas", x = 31.7, y = -31.7,
    #          size = 5, angle = 50, colour = "black") +
  guides(fill = guide_colourbar()) +
  scale_x_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°E", sep = "")) +
    scale_y_continuous(expand = c(0, 0),
                       labels = scales::unit_format(unit = "°S", sep = "")) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 10),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5),
        axis.text = element_text(size = 10)) +
    guides(fill = guide_colourbar(barheight = 1.00, barwidth = 10)) +
  theme(panel.background = element_rect(fill = "ivory", colour = NA),
        panel.border = element_rect(colour = "black", size = 0.5),
        panel.grid.minor = element_line(colour = "NA"),
        panel.grid.major = element_line(colour = "ivory", size = 0.2, linetype = "dotted"),
        legend.direction = "horizontal",
        legend.justification = c(1,0),
        legend.position = c(0.85, 0.70),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.ticks = element_line(size = 0.5))
site_map

@@

vp1 <- viewport(x = 0.70, y = 0.58, w = 0.25, h = 0.25) # Africa
vp2 <- viewport(x = 1.0, y = 1.0, w = 1.00, h = 1.00, just = c("right", "top"))  # South Africa
print(site_map, vp = vp2)
print(africa_map, vp = vp1)
dev.off()
pdf("Figures/map_fixed.pdf", width = 10, height = 7, pointsize = 4) # Set PDF dimensions
